---
title: "Military Surplus Import & Clean"
output: html_notebook
---


## Goals of This Notebook

- Download the data
- Import the data
- Check datatypes
- Create a total_value variable
- Create a control_type variable
- Filter to Texas agencies
- Filter the date range (since Jan. 1 2010)
- Export the cleaned data


## Setup

janitor is not necessary here because data already has clean names

```{r setup}
library(tidyverse)
library(DT)
```

## Download the Data

While the data I will use here is from Professor Christian McDonald, it is originally from the [Law Enforcement Support Office](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/PublicInformation/). Find more information [about the program here](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/ProgramFAQs/).


Comment out line once downloaded, so that it doesn't keep running.


```{r download}
#download.file("https://raw.githubusercontent.com/utdata/rwdir/main/data-raw/leso.csv", "data-raw/leso.csv")
```


## Import data


```{r import}
leso <- read_csv("data-raw/leso.csv")

leso
```
# Glimpse the data


```{r glimpse}
leso |> 
  glimpse()
```


## Remove unnecessary columms


```{r remove-columns}
leso_tight <- leso |> 
  select(-sheet)

leso_tight |> glimpse()
```


## Create total value column

Data only has acquisition value, which is cost per unit. Need to multiply acquisition value by quantity and create new total column. 

```{r total-value}
leso_total <- leso_tight |> 
  mutate(total_value = acquisition_value * quantity)

leso_total |> glimpse()
```

## Controlled vs. non-controlled columns & categorization logic

Need to separate the controlled vs. non-controlled items and can do this by whether or not it's an airplane or by its code. 

Goals: 

- Create a new column to denote if the item is controlled.
- In that column, needs to be TRUE when an item is controlled and FALSE when it is not.
- Items with “AIRPLANE” are always controlled, no matter their demil designations.
- Otherwise, items that have a demil_code of “A”, OR a demil_code of “Q” AND a demil_id of “6”, are non-controlled.
- Everything else is controlled.

```{r controlled}
leso_control <- leso_total |> 
  mutate(
    control_type = case_when(
      str_detect(item_name, "AIRPLANE") ~ TRUE, (demil_code == "A" | (demil_code == "Q" & demil_ic == 6)) ~ FALSE, TRUE ~ TRUE
    )
  )

leso_control |> glimpse()
```

# Confirm whether tests worked

```{r control-test}
leso_control |> 
  select(item_name, demil_code, demil_ic, control_type) |> 
  filter(str_detect(item_name, "AIRPLANE"))
```

```{r}
leso_control |> 
  count(demil_code, demil_ic, control_type, name = "cnt") |> 
  pivot_wider(names_from = control_type, values_from = cnt) |> 
  DT::datatable()
```


## Finally, filter the data

Filter the data to get only Texas observations of a certain date range. 

```{r filter}
leso_filtered <- leso_control |> 
  filter(state == "TX", ship_date >= "2010-01-01")

leso_filtered |> summary()
```


## Export the cleaned data


```{r export}
leso_filtered |> write_rds("data-processed/01-leso-tx.rds")
```

