---
title: "Military Surplus Analysis"
output: html_notebook
---

## Goals and Questions to Answer

All answers should be based on data about Texas agencies from Jan. 1, 2010 to present. All of these questions are for the “controlled” items, only.

- How many total “controlled” items were transferred, and what are they all worth? Summarize all the controlled items only to get the total quantity and total value of everything.
- How many total “controlled” items did each agency get and how much was it all worth? Which agency got the most stuff?
- How about local agencies? 
- What specific “controlled” items did each agency get and how much were they worth? Look at the kinds of items.
- What did local agencies get?


## Setup

```{r setup}
library(tidyverse)
```


## Import Cleaned Data

```{r import}
tx <- read_rds("data-processed/01-leso-tx.rds")

tx |> glimpse()
```


## Filter for only controlled items

```{r control-filter}
tx_c <- tx |> 
  filter(control_type == TRUE)

tx_c |> glimpse()
```


## Building summaries with math

Looking for how many total controlled items were transferred and what the total value was across the board.

```{r totals}
tx_c |> 
  summarise(
    summed_quantity = sum(quantity),
    summed_total_value = sum(total_value)
  )
```


## Totals by agency

For each agency, how many things did they get and how much was it all worth?

```{r agency-totals}
# save the pre-existing "by agency" list into a new object for future use.
tx_agency_totals <- tx_c |> 
  group_by(agency_name) |> 
  summarise(
    summed_quanity = sum(quantity),
    summed_total_value = sum(total_value)
  ) |> 
  arrange(desc(summed_total_value))

# peek at result
tx_agency_totals
```

## Filter for local agencies

Create a vector (or list) of the local agencies and save that into an R object for convenience. Then use the `%in%` operator to filter within a vector.

```{r local-agencies}
local_agencies <- c(
  "AUSTIN PARKS POLICE DEPT", #NI
  "AUSTIN POLICE DEPT",
  "BASTROP COUNTY SHERIFF'S OFFICE",
  "BASTROP POLICE DEPT",
  "BEE CAVE POLICE DEPT",
  "BUDA POLICE DEPT",
  "CALDWELL COUNTY SHERIFFS OFFICE",
  "CEDAR PARK POLICE DEPT",
  "ELGIN POLICE DEPARTMENT",
  "FLORENCE POLICE DEPT", #NI
  "GEORGETOWN POLICE DEPT",
  "GRANGER POLICE DEPT", #NI
  "HAYS CO CONSTABLE PRECINCT 4",
  "HAYS COUNTY SHERIFFS OFFICE",
  "HUTTO POLICE DEPT",
  "JARRELL POLICE DEPT", #NI
  "JONESTOWN POLICE DEPT", #NI
  "KYLE POLICE DEPT",
  "LAGO VISTA POLICE DEPT",
  "LAKEWAY POLICE DEPT", #NI
  "LEANDER POLICE DEPT",
  "LIBERTY HILL POLICE DEPT", #NI
  "LOCKHART POLICE DEPT",
  "LULING POLICE DEPT",
  "MANOR POLICE DEPT",
  "MARTINDALE POLICE DEPT", #NI
  "PFLUGERVILLE POLICE DEPT",
  "ROLLINGWOOD POLICE DEPT", #NI
  "SAN MARCOS POLICE DEPT",
  "SMITHVILLE POLICE DEPT", #NI
  "SUNSET VALLEY POLICE DEPT", #NI
  "TAYLOR POLICE DEPT", #NI
  "THRALL POLICE DEPT", #NI
  # TEXAS STATE UNIVERSITY HI_ED
  "TRAVIS COUNTY SHERIFFS OFFICE",
  # TRAVIS CONSTABLE OFFICE,
  # SOUTHWESTERN UNIVERSITY HI_ID
  "WESTLAKE HILLS POLICE DEPT", #NI
  "UNIV OF TEXAS SYSTEM POLICE HI_ED",
  "WILLIAMSON COUNTY SHERIFF'S OFFICE"
)

tx_agency_totals |> 
  filter(agency_name %in% local_agencies) 
  
```


## Types of items shipped to each agency

What specific controlled items did each agency get and how much were they worth? Need to group by both agency name and item name

```{r}
tx_agency_item_totals <- tx_c |> 
  group_by(agency_name, item_name) |> 
  summarise(
    summed_quantity = sum(quantity),
    summed_total_value = sum(total_value)
  ) |> 
  arrange(desc(summed_total_value))

tx_agency_item_totals
```


## Filter to get items for local agencies

```{r}
tx_agency_item_totals |> 
  filter(agency_name %in% local_agencies) |> 
  arrange(agency_name, desc(summed_total_value))
```


