---
title: "Military Surplus Data Viz"
output: html_notebook
---

## Setup

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
```

## Import the leso data

```{r import}
leso <- read_csv("data-raw/leso.csv")

leso |> glimpse()
```


## Prepare data for visualization

Need to do two things here:

- Remove unnecessary columns using select()
- Create a total_value column using mutate()

```{r clean}
leso_total <- leso |>
  select(
    -sheet,
    -nsn,
    -starts_with("demil")
  ) |> 
  mutate(total_value = quantity * acquisition_value)

leso_total |> glimpse()
```

## Wrangle the data

Goals here:

- Filter the data to focus on more recent data. Consider military surplus in 2010 and after.
- Create a year variable using the `year()` function.
- Select a few variables to study (specifically, year, total_value, and state)
- Save all that into a new data frame called leso_dated

```{r all-states}
leso_dated <- leso_total |> 
  filter(ship_date >= "2010-01-01") |> 
  mutate(year = year(ship_date)) |> 
  select(year, total_value, state)

leso_dated |> glimpse()
```

```{r texas}
leso_tx_dated <- leso_total |> 
  filter(state == "TX", ship_date >= "2010-01-01") |> 
  mutate(year = year(ship_date)) |> 
  select(year, total_value, state)

leso_tx_dated |> glimpse()
```


## Get total values by year

Group by year, add up total values and arrange in chronological order

```{r tx-totals}
leso_tx_total <- leso_tx_dated |> 
  group_by(year) |> 
  summarise(total_yearly_value = sum(total_value)) |> 
  mutate(total_yearly_value = total_yearly_value/1000000) |> # divide by million to get around notation
  arrange(year)

leso_tx_total
```


## Create a line chart for Texas data

```{r line-plot}
tx_plot <- ggplot(leso_tx_total, aes(x = year, y = total_yearly_value)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Yearly value of military surplus acquisitions in Texas",
    caption = "Source: Law Enforcement Support Office",
    x = "",
    y = "Cost of acquisitions (in millions)"
  ) +
  scale_x_continuous(breaks = seq(from = 2010, to = 2022, by = 2)) +
  theme(plot.title.position = "plot")

tx_plot
```


## Prepare data for neighboring states comparison

Want to see how Texas' military surplus info compares to that of neighboring states. Need to group by state AND year.

```{r}
leso_five <- leso_dated |> 
  filter(
    state %in% c("TX", "OK", "AR", "NM", "LA")
  ) |> 
  group_by(state, year) |> 
  summarise(total_yearly_value = sum(total_value)) |> 
  mutate(total_yearly_value = total_yearly_value/1000000)

leso_five |> glimpse()
```


## Plot multiple line chart

```{r multiple-line-chart}
five_plot <- ggplot(leso_five, aes(x = year, y = total_yearly_value)) +
  geom_point(aes(color = state)) +
  geom_line(aes(color = state)) + # the aes selects a color for each state
  labs(
    title = "Yearly Value of Military Surplus Acquisitions in Texas and Neighboring States",
    caption = "Chart: Cecilia Garzella • Source: Law Enforcement Support Office",
    x = "",
    y = "Cost of acquisitions (in millions)"
  ) +
  scale_x_continuous(breaks = seq(from = 2010, to = 2022, by = 2)) +
  theme(plot.title.position = "plot")
  #theme_minimal()

five_plot
```


## Facet wrap

```{r facet-wrap}
facet_wrap_five <- five_plot +
  facet_wrap(~ state, nrow = 2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + #rotates angle of x-axis ticks
  theme(panel.spacing.x = unit(1, "lines")) + # increases horizontal space between faceted plots
  theme(legend.position = "none") # this line removes the legend

facet_wrap_five
```



## Interactive plot

```{r interactive}
library(plotly)

tx_plot |> 
  ggplotly()
```


## Save plot

```{r}
ggsave("images/multiple-line-five.png", plot = five_plot)
```


